Spatial Reasoning and Statistical Graphics

### 2014-10-22 Paper/Draft1/TestingVisualAptitude.Rnw
Fix reference

\section{Introduction}
\comment{Intro about statistical graphics, lineups, ...}
\subsection{Spatial Reasoning and Statistical Graphics}
Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers. Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph. Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli). Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}. Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements. Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \citep{lowrie2007solving}.

>>>>

\section{Introduction}
\comment{Intro about statistical graphics, lineups, ...}
\subsection{Spatial Reasoning and Statistical Graphics}
Statistical graphics provide quick summaries of data, models, and results, but these displays may not be equally useful to all viewers. Mathematical ability is important, even for the simplest graphs: \citet{shah1995conceptual} showed that mathematical ability, but not spatial ability, was associated with accuracy on a simple two-dimensional line graph. Spatial ability becomes more important, however, when more complicated graphical displays are used in comparison tasks: \citet{mayer1994whom} hypothesize that those with low spatial ability perform worse on tests utilizing diagrams and graphs because more cognitive resources are required to process the visual stimuli (leaving fewer resources to make connections and draw conclusions from those stimuli). Many theories of graphical learning center around the difference between visual and verbal processing: the dual-coding theory emphasizes the utility of complementary information in both domains, while the visual argument hypothesis emphasizes that graphics are more efficient tools for providing data with spatial, temporal, or other implicit ordering, because the spatial dimension can be represented graphically in a more natural manner \citep{vekiri2002value}. Both of these theories suggest spatial ability would impact a viewer's use of graphics, because spatial ability either influences cognitive resource allocation or affects the processing of spatial relationships between graphical elements. Previous investigations into graphical learning and spatial ability have found relationships between spatial ability and the ability to read information from graphs \citep{lowrie2007solving}.
